---
title: "Programming Background"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "../images/")
```

```{r render, echo=FALSE, eval=FALSE}
rmarkdown::render("2021-9-12-Second-Blog-Post.Rmd",
          output_format = "github_document",
          output_dir = "../_posts/",
          output_options = list(
            html_preview = FALSE
            )
          )
```

## Comparing R to other programming languages

_What are your thoughts on R vs. whatever other software you've used? What functionality do you like about R? What parts do you miss about your other language? Do you consider R a difficult language to learn?_

After my brief forays into Base SAS in ST555 and ST513, I find that the syntax of R has been more lexically intuitive and easier to recall. Importing data, subsetting it, and converting it from a wide to long format in R has (so-far) required a one-line function, whereas one must recall a paragraph of syntax for the various SAS "DATA" and "PROC" steps. I may have overlooked the equivalent of RStudio's `help()` function in SAS, but I distinctly remember perusing lengthy SAS documentation web pages for details that are otherwise at my fingertips in RStudio. Conversely, SAS offered a relatively consistent way to perform basic analyses, such as finding the five-number summary (e.g. via `PROC MEANS DATA=... min p25 p50 p75 max`). The sheer diversity of packages in R means that, depending on which guide I read, I may use functions from `dplyr`, `stats`, `mosaic`, or potentially a host of other sources. 

Creating simple files with R Markdown and `knitr` has also been far easier than with the SAS Output Delivery System (ODS). On the other hand, SAS output was highly customizable. I recall modifying cell widths, text size and alignment, background colors, and a variety of plot properties in ST555, but I have stuck with defaults while working with R. This may change when we explore `ggplot2`.  

Lastly, our default development environments for SAS and R - SAS Studio and RStudio, respectively - have been phenomenal on the Windows and Mac operating systems, but the `tidyverse` package was surprisingly difficult to install in RStudio on Ubuntu. In the end, I resorted to an [online blog post](https://blog.zenggyu.com/en/post/2018-01-29/installing-r-r-packages-e-g-tidyverse-and-rstudio-on-ubuntu-linux/) to identify and install a slew of missing dependencies. Now that the clud-based SAS OnDemand for Academics has completely replaced SAS University Edition, getting started with SAS is, by comparison, a more seamless process on Linux. The only drawback I have faced with SAS OnDemand for Academics is its integration with a local installation of Jupyter, which requires users to create [several configuration files](https://support.sas.com/ondemand/saspy.html). The Anaconda ecosystem, on the other hand, runs identically on Windows, Mac, and Linux and provides [out-of-the-box support](https://docs.anaconda.com/anaconda/navigator/tutorials/r-lang/) for the R kernel in Jupyter.        


## Example R Markdown output

```{r pressure, echo=TRUE}
# Monthly totals of international airline passengers, 1949 to 1960.
plot(AirPassengers)
```

